### 2021-12-19 newvariant.tex
tweak

Inferring the relative strength (i.e., the ratio of reproduction numbers) and relative speed (i.e., the difference between growth rates) of new SARS-CoV-2 variants compared to earlier strains is critical to predicting and controlling the course of the current pandemic.
Analyses of new variants have primarily focused on characterizing changes in the \emph{proportion} of new variants, implicitly assuming that the relative speed remains fixed over the course of an invasion.
We use the generation-interval-based framework to challenge this assumption and illustrate how relative strength and speed change over time under two idealized interventions:
a constant-strength intervention like idealized vaccination and social distancing, which reduces transmission rates by a constant amount, and a constant-speed intervention like idealized contact tracing, which isolates infected individuals at a constant rate.
A constant-strength intervention causes relative speed to change, whereas a constant-speed intervention causes relative strength to change.
Differences in the generation-interval distributions between variants can exaggerate these changes and affect the effectiveness of interventions.
Neglecting the differences in the generation-interval distributions differences gives biased estimates of relative strength, which can vary systematically with changes in the incidence of infections caused by the variant.

>>>>

Inferring the relative strength (i.e., the ratio of reproduction numbers) and relative speed (i.e., the difference between growth rates) of new SARS-CoV-2 variants compared to earlier strains is critical to predicting and controlling the course of the current pandemic.
Analyses of new variants have primarily focused on characterizing changes in the \emph{proportion} of new variants, implicitly or explicitly assuming that the relative speed remains fixed over the course of an invasion.
We use the generation-interval-based framework to challenge this assumption and illustrate how relative strength and speed change over time under two idealized interventions:
a constant-strength intervention like idealized vaccination and social distancing, which reduces transmission rates by a constant amount, and a constant-speed intervention like idealized contact tracing, which isolates infected individuals at a constant rate.
A constant-strength intervention causes relative speed to change, whereas a constant-speed intervention causes relative strength to change.
Differences in the generation-interval distributions between variants can exaggerate these changes and affect the effectiveness of interventions.
Neglecting the differences in the generation-interval distributions differences gives biased estimates of relative strength, which can vary systematically with changes in the incidence of infections caused by the variant.
